Carcinoma

### 2022-05-15 inst/carcinoma.Rmd
carcinoma analysis

@@

for(i in seq_len(nrow(carcinoma_table))){
      jj <- unlist(carcinoma_table[i,1:7])
      negative <- names(jj)[jj==0]
      positive <- names(jj)[jj==1]
      print(negative)
      print(positive)
      Wi[[i]] <- ggrl(W,positive,negative)
}
HL <- lsl(Wi,powers= carcinoma_table$n)
```

Now

```{r,label=calccloglikzipf,cache=TRUE}
z <- zipf(7)
names(z) <- LETTERS[1:7]
system.time(lz <- loglik_lsl(z,HL))
lz
```

```{r,label-calcloglikequalp,cache=TRUE}
e <- equalp(W)
system.time(le <- loglik_lsl(e,HL))
le
```

@@

```{r}
carcinoma <- 5

```{r,label=savecarcinomadataset}
save(carcinoma_table,carcinoma, file="carcinoma.rda")
```

>>>>

This is a short analysis of the carcinoma dataset presented by
Agresti, table 13.1, p542.  Seven pathologists classified each of 118
slides on the presence or absence of carcinoma in the uterine
cervix.

The probability model is as follows.  To each pathologist $i,1\leq i
\leq 7$, we assign a non-negative Bradley-Terry strength and require
$\sum p_i=1$.  Given that two pathologists $i$ and $j$ differ in their
diagnosis then the probability of $i$ diagnosing the presence, and $j$
diagnosing the absence, of carcinoma is given by

\begin{equation}\tag{1}
i^+,j^-\frac{p_i}{p_i+p_j},\qquad i^-,j^+\frac{p_j}{p_i+p_j}
\end{equation}

Operationally the way we implement this in `hyper2` idiom is to posit
an (unobservable) ranking on the clinicians: we place the clinicians
in order from most convinced of the presence of carcinoma to the least
convinced.  Thus a ranking of `DACBEGF` means that clinician `D` was
most convinced of the presense of carcinoma and `F` least convinced.
Given this, and given a single slide, our actual observation might be
that `D` and `A` diagnosed presence, and the others diagosed absence,
of carcinoma.  The corresponding entry would be

```
A B C D E F G
T F F T F F F
```

We ask what rankings are consistent with this observation.  Any
ranking with `A` and `D` ahead of `BCEFG` would be admissible.  In
`hyper2` notation we would say

$$\left\lbrace A,D\right\rbrace\succ\left\lbrace B,C,E,F,G\right\rbrace$$

Any ranking where `A` and `D` occupy the first two ranks and
`B`,`C`,`E`,`F`,`G` ranks three to seven, would be admissible.  We
would have $2!5!=240$ rankings.  The probability of the actual ranking
being one of these is simply the sum of these 48 as the rankings are
mutually exclusive.

Each ranking has a Plackett-Luce likelihood function which is created
in `hyper2` idiom using the bespoke `race()` function.  Here, we use
generalized grouped rank likelihood (`ggrl()`) to create an `lsl`
object, as used by Hankin 2019 for the MasterChef analysis.  This
furnishes a Bradley-Terry model for the clinicians' strength that
recovers equation (1).

@@

for(i in seq_len(nrow(carcinoma_table))){
      jj <- unlist(carcinoma_table[i,1:7])
      negative <- names(jj)[jj==0]
      positive <- names(jj)[jj==1]
      print(negative)
      print(positive)
      Wi[[i]] <- ggrl(W,positive,negative)
}
carcinoma <- lsl(Wi,powers= carcinoma_table$n)
```

To maximize the likelihood we need a good starting point and rather
than use the default `equalp()` we could simply count how many times
each clinician diagnoses carcinoma:

```{r,label=countcarcinoma}
count <- colSums(sweep(carcinoma_table[,1:7],1,carcinoma_table$n,"*"))
names(count) <- colnames(carcinoma_table)[1:7]
count <- count/sum(count)
count
pie(count)
carcinoma_count <- count
```

We can compare this to other reasonable starting points:

```{r,label=calccloglikcount,cache=TRUE}
system.time(lc <- loglik_lsl(count,carcinoma))
```

```{r,label=calccloglikzipf,cache=TRUE}
z <- zipf(7)
names(z) <- LETTERS[1:7]
system.time(lz <- loglik_lsl(z,carcinoma))
```

```{r,label-calcloglikequalp,cache=TRUE}
e <- equalp(W)
system.time(le <- loglik_lsl(e,carcinoma))
```

```{r,label=showthreemethods}
c(count=lc, zipf=lz, equal=le)
```

To find the maximum likelihood estimate we would use:

`maxp_lsl(carcinoma,startp=indep(carcinoma_count), control=list(trace=100))`

but this will take many hours (a week?) to run.

@@

```{r,label=savecarcinomadataset}
save(carcinoma_table,carcinoma, carcinoma_count, file="carcinoma.rda")
```
